---
title: "ENERGY REQUIREMENTS OF GREY WHALES"
author: "Selina Agbayani"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  html_document:
 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(warning = FALSE)

## Setup

# Initialize all libraries, set paths for output figures, then import the gw_observations.csv
# install.packages("tidyr")
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("ggplot2")
# install.packages("knitr")
# install.packages("nls2")
# install.packages("nlstools")
# install.packages("nlme")
# install.packages("minpack.lm")
# install.packages("scales")
# install.packages("ggpmisc")
# install.packages("extrafont")
#install.packages("truncnorm")



library(tidyverse)
# library(tidyr)
# library(dplyr)
# library(ggplot2)
library(knitr)
# library(nls2)
# library(nlstools)
# library(nlme)
# library(minpack.lm)
library(scales)
library(ggpmisc)
library(reshape2)
library(extrafont)
library(truncnorm)
library(gtable)
library(ggpubr)

#source("functions.R")

#quiet col types messages
options(readr.show_col_types = FALSE)

```


```{r declare functions and global variables, include=FALSE}

# **Declare custom functions and global variables**
########## Specify Decimal function ############################
# 
# For use in labelling the plots with equations, and rounding up the coefficients
# to a specific no. of decimal places
specify_decimal <- function(x, k) format(round(x, k), nsmall=k)


############ Multiple plot function######################
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
# Code extracted from Cookbook for R. This site is powered by knitr and Jekyll. 
# If you find any errors, please email winston@stdout.org

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}


########### functions used for carry out the fit ########

## this function is avaible in: 
# http://www.leg.ufpr.br/~walmes/cursoR/ciaeear/as.lm.R   or in 
# https://gist.github.com/TonyLadson/2d63ca70eef92583001dece607127759 (from the line 269)

##### as.lm function: ######

        as.lm <- function(object, ...) UseMethod("as.lm")

        as.lm.nls <- function(object, ...) {
            if (!inherits(object, "nls")) {
                w <- paste("expected object of class nls but got object of class:", 
                    paste(class(object), collapse = " "))
                warning(w)
            }

            gradient <- object$m$gradient()
            if (is.null(colnames(gradient))) {
                colnames(gradient) <- names(object$m$getPars())
            }

            response.name <- if (length(formula(object)) == 2) "0" else 
                as.character(formula(object)[[2]])

            lhs <- object$m$lhs()
            L <- data.frame(lhs, gradient)
            names(L)[1] <- response.name

            fo <- sprintf("%s ~ %s - 1", response.name, 
                paste(colnames(gradient), collapse = "+"))
            fo <- as.formula(fo, env = as.proto.list(L))

            do.call("lm", list(fo, offset = substitute(fitted(object))))

        }

############## End as.lm function 



############### proto function: ############
#### proto function avaible in https://github.com/hadley/proto/blob/master/R/proto.R

proto <- function(. = parent.env(envir), expr = {},
                   envir = new.env(parent = parent.frame()), ...,
                   funEnvir = envir) {
  parent.env(envir) <- .
  envir <- as.proto.environment(envir)  # must do this before eval(...)
  # moved eval after for so that ... always done first
  # eval(substitute(eval(quote({ expr }))), envir)
  dots <- list(...); names <- names(dots)
  for (i in seq_along(dots)) {
    assign(names[i], dots[[i]], envir = envir)
    if (!identical(funEnvir, FALSE) && is.function(dots[[i]]))
      environment(envir[[names[i]]]) <- funEnvir
  }
  eval(substitute(eval(quote({
    expr
  }))), envir)
  if (length(dots))
    as.proto.environment(envir)
  else
    envir
}

#' @export
#' @rdname proto
as.proto <- function(x, ...) {
  UseMethod("as.proto")
}

#' @export
#' @rdname proto
as.proto.environment <- function(x, ...) {
  assign(".that", x, envir = x)
  assign(".super", parent.env(x), envir = x)
  structure(x, class = c("proto", "environment"))
}

#' @export
#' @rdname proto
as.proto.proto <- function(x, ...) {
  x
}
as.proto.list <- function(x, envir, parent, all.names = FALSE, ...,
                          funEnvir = envir, SELECT = function(x) TRUE) {
  if (missing(envir)) {
    if (missing(parent))
      parent <- parent.frame()
    envir <- if (is.proto(parent))
      parent$proto(...)
    else
      proto(parent, ...)
  }
  for (s in names(x))
    if (SELECT(x[[s]])) {
      assign(s, x[[s]], envir = envir)
      if (is.function(x[[s]]) && !identical(funEnvir, FALSE))
        environment(envir[[s]]) <- funEnvir
    }
  if (!missing(parent))
    parent.env(envir) <- parent
  as.proto.environment(envir)  # force refresh of .that and .super
}

#' @export
"$<-.proto" <- function(this,s,value) {
  if (s == ".super")
    parent.env(this) <- value
  if (is.function(value))
    environment(value) <- this
  this[[as.character(substitute(s))]] <- value
  this
}
is.proto <- function(x) inherits(x, "proto")

#' @export
"$.proto" <- function(x, name) {
  inherits <- substr(name, 1, 2) != ".."

  res <- get(name, envir = x, inherits = inherits)
  if (!is.function(res))
    return(res)

  if (deparse(substitute(x)) %in% c(".that", ".super"))
    return(res)

  structure(
    function(...) res(x, ...),
    class = "protoMethod",
    method = res
  )
}

#' @export
print.protoMethod <- function(x, ...) {
  cat("<ProtoMethod>\n")
  print(attr(x, "method"), ...)
}

# modified from Tom Short's original
#' @export
str.proto <- function(object, max.level = 1, nest.lev = 0,
                      indent.str = paste(rep.int(" ", max(0, nest.lev + 1)), collapse = ".."),
                      ...) {
  cat("proto", name.proto(object), "\n")
  Lines <- utils::capture.output(utils::str(
    as.list(object), max.level = max.level,
    nest.lev = nest.lev, ...
  ))[-1]
  for (s in Lines)
    cat(s, "\n")
  if (is.proto(parent.env(object))) {
    cat(indent.str, "parent: ", sep = "")
    utils::str(parent.env(object), nest.lev = nest.lev + 1, ...)
  }
}

#' @export
print.proto <- function(x, ...) {
  if (!exists("proto_print", envir = x, inherits = TRUE))
    return(NextMethod())

  x$proto_print(...)
}

############# End proto function 

############ PredictNLS - for confidence intervals  ##########
# Source: https://rmazing.wordpress.com/2013/08/14/predictnls-part-1-monte-carlo-simulation-confidence-intervals-for-nls-models/

predictNLS <- function(object,newdata,level = 0.95, nsim = 10000,
                       ...){
    require(MASS, quietly = TRUE)
    ## get right-hand side of formula
    RHS <- as.list(object$call$formula)[[3]]
    EXPR <- as.expression(RHS)
   
    ## all variables in model
    VARS <- all.vars(EXPR)
   
    ## coefficients
    COEF <- coef(object)
   
    ## extract predictor variable    
    predNAME <- setdiff(VARS, names(COEF))  
   
    ## take fitted values, if 'newdata' is missing
    if (missing(newdata)) {
        newdata <- eval(object$data)[predNAME]
        colnames(newdata) <- predNAME
        }
       
    ## check that 'newdata' has same name as predVAR
    if (names(newdata)[1] != predNAME) stop("newdata should have name '",
                                            predNAME, "'!")
   
    ## get parameter coefficients
    COEF <- coef(object)
     
    ## get variance-covariance matrix
    VCOV <- vcov(object)
   
    ## augment variance-covariance matrix for 'mvrnorm' 
    ## by adding a column/row for 'error in x'
    NCOL <- ncol(VCOV)
    ADD1 <- c(rep(0, NCOL))
    ADD1 <- matrix(ADD1, ncol = 1)
    colnames(ADD1) <- predNAME
    VCOV <- cbind(VCOV, ADD1)
    ADD2 <- c(rep(0, NCOL + 1))
    ADD2 <- matrix(ADD2, nrow = 1)
    rownames(ADD2) <- predNAME
    VCOV <- rbind(VCOV, ADD2) 
         
    ## iterate over all entries in 'newdata' as in usual 'predict.' functions
    NR <- nrow(newdata)
    respVEC <- numeric(NR)
    seVEC <- numeric(NR)
    varPLACE <- ncol(VCOV)   
   
    ## define counter function
    counter <- function (i) {
        if (i%%10 == 0) 
            cat(i)
        else cat(".")
        if (i%%50 == 0) 
            cat("\n")
        flush.console()
        }
   
    outMAT <- NULL 
   
    for (i in 1:NR) {
        counter(i)
        ## get predictor values and optional errors
    predVAL <- newdata[i, 1]
    if (ncol(newdata) == 2) predERROR <- newdata[i, 2] else predERROR <- 0
    names(predVAL) <- predNAME  
    names(predERROR) <- predNAME  
     
    ## create mean vector for 'mvrnorm'
    MU <- c(COEF, predVAL)
     
    ## create variance-covariance matrix for 'mvrnorm'
    ## by putting error^2 in lower-right position of VCOV
    newVCOV <- VCOV
    newVCOV[varPLACE, varPLACE] <- predERROR^2
     
    ## create MC simulation matrix
    simMAT <- mvrnorm(n = nsim, mu = MU, Sigma = newVCOV, empirical = TRUE)
     
    ## evaluate expression on rows of simMAT
    EVAL <- try(eval(EXPR, envir = as.data.frame(simMAT)), silent = TRUE)
    if (inherits(EVAL, "try-error")) stop("There was an error evaluating the simulations!")
     
    ## collect statistics
    PRED <- data.frame(predVAL)
    colnames(PRED) <- predNAME   
    FITTED <- predict(object, newdata = data.frame(PRED))
    MEAN.sim <- mean(EVAL, na.rm = TRUE)
    SD.sim <- sd(EVAL, na.rm = TRUE)
    MEDIAN.sim <- median(EVAL, na.rm = TRUE)
    MAD.sim <- mad(EVAL, na.rm = TRUE)  #median absolute deviation
    QUANT <- quantile(EVAL, c((1 - level)/2, level + (1 - level)/2))
    RES <- c(FITTED, MEAN.sim, SD.sim, MEDIAN.sim, MAD.sim, QUANT[1], QUANT[2])
    outMAT <- rbind(outMAT, RES)
  }
   
  colnames(outMAT) <- c("fit", "mean", "sd", "median", "mad", names(QUANT[1]), names(QUANT[2]))
  rownames(outMAT) <- NULL
   
  cat("\n")
   
  return(outMAT)  
}

##########end PredictNLS ##########

        
```   



```{r setpaths, include = FALSE}
# Set path for output figures: 
Figurespath <- paste0(getwd(), "/FMR/figures", collapse = NULL)
Figurespath
# Set path for input & output data  
datapath <- paste0(getwd(), "/data", collapse = NULL) 
datapath

```

```{r name_your_chunk, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

```

## RESULTS

### Production cost (P)

#### Grey whale calves

```{r P_cost_phase1, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

P_cost_table  <-  read_csv("data/P_cost_table_phase1.csv")

plot_P_cost_mass_perday <- P_cost_table %>%
  dplyr::filter(sex == "N/A") %>%
  ggplot() +
  # geom_errorbar(aes(x = age_mth-0.5, 
  #                   ymin = (mass - mass_sd) /50, 
  #                 ymax = (mass + mass_sd) /50), 
  #               width=0, size=1, linetype = 2, color="gray40") +
  geom_ribbon(aes(x = age_mth, 
                  ymin = (mass - mass_sd) /12, 
                  ymax = (mass + mass_sd) /12), 
              fill = "gray", alpha = 0.8)+
  geom_line(aes(x = age_mth, y = mass/12), 
             linetype = 2, size = 0.5, colour = "black")+
  geom_errorbar(aes(x = age_mth-0.5, 
                    ymin = mean_P_perday - sd_P_perday, 
                    ymax = mean_P_perday + sd_P_perday), 
                width=0, size=0.5, linetype = 1, color="gray40") +
  geom_point(aes(x = age_mth-0.5, y= mean_P_perday), size = 2)+
  xlab("Age (months)") +
  ylab(bquote('Mean production cost (P) in MJ day'^'-1')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10),
                     limits = c(0, 12)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),
                     limits = c(0, 550),
                     labels = scales::comma,
                     sec.axis=sec_axis(~.*12, name = "Mass (kg)",
                                       labels = scales::comma,
                                       breaks = scales::pretty_breaks(n = 10))
                     
                     )+
                     #breaks = scales::pretty_breaks(n = 8))+
                       #limits = c(0, 50)) +
  theme_bw() + 
  theme(panel.grid = element_blank())+
  theme(legend.position = "none")+
    theme(legend.background = element_rect(fill = "white"))+
    theme(legend.text = element_text(colour = "black", size = rel(1)))+
    theme(legend.key = element_rect(fill = "transparent"))+ #colour = "transparent"))+
    ggtitle(element_blank()) +
    theme(plot.background = element_rect(fill = "white"))+
    theme(panel.background = element_rect(fill = "white"))+
    theme(panel.border = element_blank())+
    theme(axis.line = element_line(colour = "black"))+
    theme(axis.text = element_text(colour = "black", size = rel(1.2)))+
    theme(axis.title.y = element_text(colour = "black", 
                                      size = rel(1.4), 
                                      angle = 90, 
                                      margin = margin(t = 0, r = 10, b = 0, l = 0)))+
    theme(axis.title.x = element_text(colour = "black", size = rel(1.4)))+
    theme(axis.ticks = element_line(colour="black"))+
    theme(axis.title.y.right  = element_text(colour = "black", 
                                      size = rel(1.2), 
                                      angle = 270, 
                                      margin = margin(t = 0, r = 0, b = 0, l = 10)))+
    annotate("text", x= 12, y = 25,
             label = "P cost", 
             hjust = 1,
             size = rel(4),
             color = "black") +
    annotate("text", x= 12, y = 550,
             label = "Mass", 
             hjust = 1,
             size = rel(4),
             color = "black")
  


plot_P_cost_mass_perday
```

#### Juveniles and Adults

```{r P_cost_peryear, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}
P_cost_table_peryear  <- read_csv("data/P_cost_table_peryear.csv")

plot_P_cost_mass_peryear <- P_cost_table_peryear %>% filter(sex == "N/A" & age_yrs >=1 & age_yrs<=30) %>%  
  ggplot() +
  geom_ribbon(aes(x = age_yrs, 
                  ymin = (mean_mass - sd_mass) /100, 
                  ymax = (mean_mass + sd_mass) /100), 
              fill = "gray", alpha = 0.8)+
  geom_line(aes(x = age_yrs, y = mean_mass/100), 
             linetype = 2, size = 0.5, colour = "black")+
  geom_ribbon(aes(x = age_yrs, 
                  ymin = (mean_P_perday - sd_P_perday),
                  ymax = (mean_P_perday + sd_P_perday)), 
                  #width = 0, linetype = 3, size = 1, colour= "white")+
              fill = "gray", alpha = 0.8)+
  geom_line(aes(x = age_yrs, y= mean_P_perday), 
              size = 0.5, linetype = 3, colour = "black")+
  xlab("Age (years)") +
  ylab(bquote('Production cost (MJ day'^'-1'*')'))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 9))+
                     #limits = c(0, 50)) +  
  scale_y_continuous(labels= scales::comma,
                     sec.axis=sec_axis(~.*100, name = "Mass (kg)",
                                       labels = scales::comma), 
                     breaks = scales::pretty_breaks(n = 8) 
                     )+
                       #limits = c(0, 50)) +
          
  theme_bw() + 
  theme(panel.grid = element_blank())+
  theme(legend.position = "none")+
    theme(legend.background = element_rect(fill = "white"))+
    theme(legend.text = element_text(colour = "black", size = rel(1)))+
    theme(legend.key = element_rect(fill = "transparent"))+ #colour = "transparent"))+
    ggtitle(element_blank()) +
    theme(plot.background = element_rect(fill = "white"))+
    theme(panel.background = element_rect(fill = "white"))+
    theme(panel.border = element_blank())+
    theme(axis.line = element_line(colour = "black"))+
    theme(axis.text = element_text(colour = "black", size = rel(1.2)))+
    theme(axis.title.y = element_text(colour = "black", 
                                      size = rel(1.4), 
                                      angle = 90, 
                                      margin = margin(t = 0, r = 10, b = 0, l = 0)))+
    theme(axis.title.x = element_text(colour = "black", size = rel(1.4)))+
    theme(axis.ticks = element_line(colour="black"))+
  theme(axis.title.y.right  = element_text(colour = "black", 
                                      size = rel(1.2), 
                                      angle = 270, 
                                      margin = margin(t = 0, r = 0, b = 0, l = 10)))+
    annotate("text", x= 30, y = 25,
             label = "Production cost", 
             hjust = 1,
             size = rel(5),
             color = "black") +
    annotate("text", x= 30, y = 220,
             label = "Mass", 
             hjust = 1,
             size = rel(5),
             color = "black") 

  
  
  


plot_P_cost_mass_peryear

```



### Total Energetic Expenditure (*E~s~*)

#### Grey whale calves

```{r Es_table_phase1_permth, echo= FALSE, results = "hide", message = FALSE, warning=FALSE }

Es_table_phase1_permth  <-  read_csv("data/Es_sensAnalysis_phase1_permth_source_bpm.csv")

plot_Es_table_phase1_mthly_perday <- Es_table_phase1_permth %>% 
  #dplyr::filter(age_yrs == i) %>% 
  ggplot(aes(x = age_mth-0.5, y =Es_perday)) +
  geom_errorbar(aes(ymin = Es_perday - Es_perday_sd, ymax = Es_perday + Es_perday_sd),
                size = 0.5, linetype = 1, colour = 'gray80', width = 0) +
  geom_point(aes(x = age_mth-0.5, y =Es_perday)) +
  xlab("Age (months)") +
  ylab(bquote('Energetic expenditure (MJ day'^'-1'*')')) +
  #ylab(bquote('GER (MJ day '^'-1'*')'))
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 11))+ 
  #limits = c(0, 30)) +  # max x-axis 30 yrs. 
  scale_y_continuous(label = comma, 
                     breaks = scales::pretty_breaks(n = 10),
                     limits = c(0,550))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.title.x = element_text(size = rel(1.4),
                                    colour = "black"))+
  theme(axis.title.y = element_text(size = rel(1.4),
                                    colour = "black"))+
  theme(axis.text = element_text(size = rel(1.2),
                                 colour = "black"))  
#ggtitle("Es table Phase 1 - per mth")

plot_Es_table_phase1_mthly_perday

```

Figure 4. Estimates of daily energetic expenditure from activity (MJ day^-1^) for each month in the first year (Phase 1). Error bars indicate SD. Note: Es values for the 6th and 12th months represent energetic expenditures across two activity stages.


#### Grey whale juveniles and adults, including pregnant and lactating females

```{r Es_table_alladults_preg_lact_peryear, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

Es_table_alladults_preg_lact_peryear  <- 
  read_csv("data/Es_sensAnalysis_alladults_preg_lact_peryear_source_bpm.csv")

# apply factors to enable facet wrap 
Es_table_alladults_preg_lact_peryear$Lifestage_f = factor(Es_table_alladults_preg_lact_peryear$Lifestage,
                                                          levels=c('Juvenile/Adult','Pregnant','Lactating'))


# plots  

plot_Es_table_alladults_preg_lact_yearly_perday <- 
  Es_table_alladults_preg_lact_peryear %>% 
  ggplot(aes(x = age_yrs, y =Es_perday)) + #shape = Lifestage
  facet_grid(. ~Lifestage_f)+
  geom_errorbar(aes(x = age_yrs, 
                    ymin = Es_perday - Es_sd_perday, 
                    ymax = Es_perday + Es_sd_perday),
                size = 0.5, colour = 'gray80', width = 0, linetype = 1) +
  # geom_ribbon(aes(x = age_yrs, ymin = Es - Es_sd, ymax = Es + Es_sd),
  #                 fill = 'gray40', alpha = 0.3) +
  geom_point()+
  #geom_line()   +
  xlab("Age (years)") +
  
  ylab(bquote('Energetic expenditure (MJ day'^'-1'*')')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), 
                     limits = c(0, 32)) +  # max x-axis 30 yrs. 
  scale_y_continuous(label = comma, 
                     breaks = scales::pretty_breaks(n = 8),
                     limits = c(0,max(Es_table_alladults_preg_lact_peryear$Es_perday +
                                        Es_table_alladults_preg_lact_peryear$Es_sd_perday)))+
  # scale_y_continuous(label = function(x){(x/1000)}, breaks = scales::pretty_breaks(n = 8),
  #                    limits = c(0, 1000000))+
  
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text = element_text(size = rel(1.2),
                                 colour = "black"))+
  theme(axis.title.x = element_text(size = rel(1.4)),
        axis.title.y = element_text(size = rel(1.2)))+
  theme(strip.background =element_rect(fill="transparent", 
                                       colour = "transparent"))+
  theme(strip.text = element_text(size = rel(1.2)))
#theme(legend.position = "none")

plot_Es_table_alladults_preg_lact_yearly_perday

```

Figure 5. Estimates of annual total energetic expenditure from activity (MJ) for juveniles and single adults, pregnant and lactating females. Error bars represent SD.

### Gross Energy Requirements

#### Grey whale calves

```{r predict_GER_table_phase1_permth, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

predict_GER_table_phase1_permth <- read_csv("data/predict_GER_table_sensAnalysis_phase1_permth_source_bpm.csv")


plot_predict_GER_phase1_permth <- predict_GER_table_phase1_permth %>%
  dplyr::filter(sex == "N/A") %>% 
  ggplot(aes(x = age_mth, y = mean_GER)) +
  geom_errorbar(aes(x = age_mth-0.5, 
                    ymin = mean_GER-GER_sd, 
                    ymax = mean_GER+GER_sd), 
                width=0, size = 0.5,
                linetype = 1,
                color="gray80") + 
  geom_point(aes(x = age_mth-0.5, y= mean_GER), 
             shape = 21, fill = "black")+
  xlab("Age (months)") +
  ylab(bquote('Gross Energy Reqs (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 12)) + 
  #limits = c(0, 30)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 7),
                     limits = c(0,1100)) + 
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text = element_text(size = rel(1.2),
                                 colour = "black")) +
  theme(axis.title.x = element_text(size = rel(1.5),
                                    colour = "black")) +
  theme(axis.title.y = element_text(size = rel(1.4),
                                    colour = "black"))



plot_predict_GER_phase1_permth


```

Figure 6. Daily Gross Energy Requirements (GER) for calves in the first year in MJ day^-1^. Error bars represent SD.

#### Grey whale juveniles / adults

```{r predict_GER_table_phase2, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

predict_GER_table_phase2 <- 
  read_csv("data/predict_GER_table_sensAnalysis_phase2.csv")

predict_GER_table_phase2 %>% 
  filter(sex=="N/A" & phase == "2" & age_yrs >= 1)%>% 
  head() %>% kable()
kable(head(predict_GER_table_phase2))

plot_predict_GER_all_yr_phase2 <- predict_GER_table_phase2 %>%
  dplyr::filter(sex=="N/A" & phase == "2" & 
                  age_yrs >= 1 & age_yrs <= 30) %>% 
  ggplot() +
  # geom_line(aes(x=age_yrs, y=mean_GER),
  #           color='black',
  #           size = 1.2,
  #           linetype = "longdash") +
  
  geom_errorbar(aes(x = age_yrs-0.5, 
                    ymin = mean_GER-GER_sd, 
                    ymax = mean_GER+GER_sd), 
                width=0, size=0.5, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_yrs-0.5, y= mean_GER), shape = 21, fill = "black")+
  xlab("Age (years)") +
  ylab(bquote('Gross Energy Requirements (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), 
                     limits = c(0, 30.5), expand = c(0,0)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),
                     limits = c(-50, 4000), expand = c(0,0),
                     labels = comma)+
  
  theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(axis.title = element_text(size = rel(1.4), colour = "black"))+
  theme(axis.text = element_text(size = rel(1.2), colour = "black"))+
  annotate("text", x = 28, y = 3500,
           label = paste0("Annual"),
           size = 6,
           hjust = 1) 

#plot_predict_GER_all_yr_phase2


plot_predict_GER_foraging_phase2 <- predict_GER_table_phase2 %>% 
  dplyr::filter(sex=="N/A" & phase == "2" & 
                  age_yrs >= 1 & age_yrs <= 30) %>% 
  ggplot() +
  geom_errorbar(aes(x = age_yrs-0.5, 
                    ymin = GER_foraging - GER_sd, 
                    ymax = GER_foraging + GER_sd), 
                width=0, size=0.5, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_yrs-0.5, y= GER_foraging), 
             shape = 21, fill = "black")+
  xlab("Age (years)") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), 
                     limits = c(0, 30.5), expand = c(0,0)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),
                     limits = c(-50, 4000), expand = c(0,0))+
  
  theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size = rel(0.2), colour = NA))+
  theme(axis.title = element_text(size = rel(1.4), colour = "black"))+
  theme(axis.text = element_text(size = rel(1.2), colour = "black"))+
  annotate("text", x = 28, y = 3500,
           label = paste0("Foraging season"),
           size = 6,
           hjust = 1) 


#plot_predict_GER_foraging_phase2


plot_predict_GER_phase2_both_annual_foraging <- 
  ggarrange(plot_predict_GER_all_yr_phase2, 
            plot_predict_GER_foraging_phase2, 
            widths = c(2, 1.8)) 
plot_predict_GER_phase2_both_annual_foraging 

```

Figure 7. Daily gross energy requirements over the year compared to gross daily requirements when foraging is limited to the summer season.  Foraging season spans \~154 days for juveniles and adults. 



```{r predict_GER_table_preg_lact, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

predict_GER_table_3panel  <- read_csv("data/predict_GER_table_all_preg_lact_source_bpm.csv")

predict_GER_table_3panel <- within(predict_GER_table_3panel, Lifestage <- 
  factor(Lifestage, levels=c('Juvenile/Adult','Pregnant', 'Lactating')))



plot_GER_3panel <- predict_GER_table_3panel %>% 
  ggplot() +
  geom_errorbar(aes(x = age_yrs-0.5, 
                  ymin = mean_GER-GER_sd,
                  ymax = mean_GER+GER_sd),
                size = 0.5, width=0, linetype = 1, color="gray80")+
  geom_point(aes(x = age_yrs-0.5, y= mean_GER),
              shape = 21, fill = "black") +
   geom_errorbar(aes(x = age_yrs-0.5, 
                    ymin = GER_foraging - sd_foraging,
                    ymax = GER_foraging + sd_foraging),
                size = 0.5, width=0, linetype = 1, color="gray80") +
  geom_point(aes(x = age_yrs-0.5, y= GER_foraging),
              shape = 21, fill = "white") +
  facet_grid(~Lifestage)+
  xlab("Age (years)") +
  ylab(bquote('Gross Energy Requirements (MJ day'^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 3),
                     limits = c(1,30)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 8),
                     label = comma)+
                     #limits = c(0, 20))+
  
  theme_bw() +
  theme(panel.grid = element_blank())+
  theme(axis.text =  element_text(size = rel(1.2), colour = "black"),
        axis.title = element_text(size = rel(1.2), colour = "black"),
        strip.background = element_rect(
          color="transparent", fill="transparent"),
        strip.text = element_text(size = rel(1.2), colour = "black"))
  # annotate("text", x = 20, y = 5200,
  #          label = paste0("Foraging season"),
  #          size = 6,
  #          hjust = 1) +
  # annotate("text", x = 20, y = 200,
  #          label = paste0("Annual"),
  #          size = 6,
  #          hjust = 1) +
  
ann_text_annual <- data.frame(age_yrs = 30, mean_GER = 900, 
                              Lifestage = factor("Juvenile/Adult",
                                           levels = c("Juvenile/Adult",
                                                      "Lactating",
                                                      "Pregnant")))

ann_text_foraging <- data.frame(age_yrs = 30, mean_GER = 3200, 
                                Lifestage = factor("Juvenile/Adult",
                                             levels = c("Juvenile/Adult",
                                                        "Lactating",
                                                        "Pregnant")))

plot_GER_3panel <- plot_GER_3panel + 
  geom_text(data = ann_text_annual, 
            aes(x=age_yrs, y=mean_GER),
            label = "Annual",
            size = 5,
            hjust = 1) +
  geom_text(data = ann_text_foraging, 
            aes(x=age_yrs, y=mean_GER),
            label = "Foraging Season",
            size = 5,
            hjust = 1)

plot_GER_3panel


```

Figure 8. Estimates for gross energy requirements for pregnant, lactating females, and others (juvenile and other adult grey whales). Daily gross energy requirements for a full year (annual) are represented with black circles. White circles represent daily gross energy requirements for when feeding is limited during the foraging season. Foraging season is assumed to be ~154 days for single juveniles (1-7 y) and adults (> 8 y), \~153 days for pregnant females, and ~119 days for lactating females. Error bars represent SD.


### Milk / Food Requirements

#### Grey whale calves

```{r plot_milk_food_reqs_phase1,  echo= FALSE, results = "hide", message = FALSE, warning=FALSE}
predict_GER_table_phase1_permth <- read_csv("data/predict_GER_table_sensAnalysis_phase1_permth_source_bpm.csv")

plot_MilkReqs_phase1 <- predict_GER_table_phase1_permth %>% 
  dplyr::filter(sex == "N/A" & phase == "1") %>% 
  ggplot() +
  #geom_line(aes(x = age_yrs, y= FR_foraging))+
  geom_errorbar(aes(x = age_mth-0.5, ymin = FR_foraging - FR_sd_foraging, ymax = FR_foraging + FR_sd_foraging),
                size = 0.5, width=0, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_mth-0.5, y= FR_foraging), shape = 21, fill = "black") +
  xlab("Age (months)") +
  ylab(bquote('Milk Requirements (L day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 9),
                     limits = c(0, 9.9), expand=c(0,0)) +  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),
                     limits = c(0, 45)) +
  theme_bw() + 
  theme(panel.grid = element_blank())+
  theme(axis.text =element_text(size = rel(1.4),
                                colour = "black"),
        axis.title.x = element_text(size = rel(1.6),
                                    colour = "black"),
        axis.title.y = element_text(size = rel(1.4),
                                    colour = "black")
  )


#plot_MilkReqs_phase1


plot_FoodReqs_phase1 <- predict_GER_table_phase1_permth %>% dplyr::filter(sex == "N/A" & phase == "1") %>% 
  ggplot() +
  #geom_line(aes(x = age_yrs, y= FR_foraging))+
  geom_errorbar(aes(x = age_mth-0.5, ymin = FR_foraging - FR_sd_foraging, ymax = FR_foraging + FR_sd_foraging),
                size = 0.5, width=0, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_mth-0.5, y= FR_foraging), 
             shape = 21, fill = "black")+
  # geom_line(aes(x = age_yrs, y= quant025), linetype = "dashed", colour = "gray20")+
  # geom_line(aes(x = age_yrs, y= quant975), linetype = "dashed", colour = "gray20")+
  #facet_wrap(~sex)+
  #xlab("Age (months)") +
  ylab(bquote('Food Requirements (kg day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 2),
                     limits = c(9.9, 12), expand=c(0,0)
  ) +  # max x-axis 30 yrs.
  scale_y_continuous(position = "right",      label=comma,
                     breaks = scales::pretty_breaks(n = 10),
                     limits = c(0, 300)
  ) +
  theme_bw()+ 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = rel(1.6),
                                    colour = "white"))+
  theme(axis.text =element_text(size = rel(1.4),
                                colour = "black"),
        axis.title.y = element_text(size = rel(1.4),
                                    colour = "black")
  )

#theme(legend.position = 0)

#plot_FoodReqs_phase1



## plot milk and food requirments for phase 1 on same chart but different axes

#multiplot(plot_MilkReqs_phase1, plot_FoodReqs_phase1, cols = 2)

milkreqs_figure <- ggarrange(plot_MilkReqs_phase1, plot_FoodReqs_phase1, widths = c(2, 0.9))
milkreqs_figure


```

Figure 9. Daily milk requirements for calves. The left panel represents milk requirements (L day^-1^) for calves up to 9 months (left panel), and the right panel represents food requirements (kg of benthic invertebrate prey day^-1^) for weaned calves over 9 months.

#### Grey whale juveniles and adults, including pregnant and lactating

```{r plot_FoodReqs_3panel, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

predict_GER_table_3panel  <- read_csv("data/predict_GER_table_all_preg_lact_source_bpm.csv")

predict_GER_table_3panel <- within(predict_GER_table_3panel, Lifestage <- 
  factor(Lifestage, levels=c('Juvenile/Adult','Pregnant', 'Lactating')))


plot_FoodReqs_3panel <- predict_GER_table_3panel %>% 
  ggplot() +
  geom_errorbar(aes(x = age_yrs, 
                  ymin = FR_foraging-FR_sd_foraging,
                  ymax = FR_foraging+FR_sd_foraging),
                size = 0.5, width=0, linetype = 1, color="gray80")+
  geom_point(aes(x = age_yrs, 
                y= FR_foraging),  
            size = 1)+
  facet_wrap(~Lifestage)+
  xlab("Age (years)") +
  ylab(bquote('Food Requirements (kg day'^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 3),
                     limits = c(1,30)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))+
                     #limits = c(0, 20))+
  
  theme_bw() +
   theme(panel.spacing = unit(0.6, "lines"))+
  theme(panel.grid = element_blank())+
  theme(axis.text =  element_text(size = rel(1.2), colour = "black"),
        axis.title = element_text(size = rel(1.2), colour = "black"),
        strip.background = element_rect(
          color="transparent", fill="transparent"),
        strip.text = element_text(size = rel(1.2), colour = "black")
  )



plot_FoodReqs_3panel



```

Figure 10. Food requirements (kg day-1) of benthic invertebrate prey for pregnant females, lactating females, and others (juveniles and other adult grey whales) during foraging season. Foraging season is assumed to be ~154 days for single juveniles (1-7 y) and adults (> 8 y), ~153 days for pregnant females, and ~119 days for lactating females. Error bars represent SD.



```{r plot_pct_bodywt_consumed_3panel, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

pct_bodyweight_all <-  read_csv("data/pct_bodyweight_all_lact_preg.csv")

pct_bodyweight_all <- within(pct_bodyweight_all, Lifestage <- 
  factor(Lifestage, levels=c('Juvenile/Adult','Pregnant', 'Lactating')))


plot_pct_bodyweight_all <- pct_bodyweight_all %>% 
  filter(age_yrs > 1) %>% 
  ggplot() +
   geom_errorbar(aes(x = age_yrs-0.5, 
                    ymin = pctbodywt - pctbodywt_sd, 
                    ymax = pctbodywt + pctbodywt_sd),
                width=0, size=0.5, linetype = 1, color="gray80")+
  geom_point(aes(x = age_yrs-0.5, y= pctbodywt), 
             shape = 21, fill = "black")+
  facet_wrap(~Lifestage)+
  xlab("Age (years)") +
  ylab(bquote('Consumption day'^'-1'*' (% body wt)')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 3),
                     limits = c(1,30)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)
                     #limits = c(3.5, 10)
                     )+
  
  theme_bw() +
  theme(panel.grid = element_blank())+
  theme(axis.text =  element_text(size = rel(1.2), colour = "black"),
        axis.title = element_text(size = rel(1.2), colour = "black"),
        strip.background = element_rect(
          color="transparent", fill="transparent"),
        strip.text = element_text(size = rel(1.2), colour = "black")
  )



plot_pct_bodyweight_all



```

Figure 11. Percentage of body weight consumed day-1 during the foraging season for juveniles and other adults, pregnant females, and lactating females. Foraging season is assumed to be ~154 days for single juveniles (1-7 y) and adults (> 8 y), ~153 days for pregnant females, and ~119 days for lactating females. Note that the y-axis does not start at 0. Error bars represent SD.


## Sensitivity Analyses

The sensitivity analyses show that the variable contributing the highest range of uncertainty (±SD) to the gross energy requirements is *E~s~* or total energetic expenditure (Figure 12), and that the variable contributing the highest range of uncertainty within the Es calculations is *%O~2~* or the percentage of oxygen absorbed (Figure 13). The pattern holds for Phase 2, pregnant, and lactating life stages, even when additional parameters are examined for the pregnant and lactating stages (see Appendix D). 

### Gross Energy Requirements (*GER*)

#### Grey whale calves

```{r predict_GER_table_sensAnalysis_phase1,  echo= FALSE, results = "hide", message = FALSE, warning=FALSE}
predict_GER_table_sensAnalysis_phase1_permth  <-  read_csv("data/predict_GER_table_sensAnalysis_phase1_permth_source_bpm.csv")


predict_GER_table_sensAnalysis_phase1_permth$MC_var_factor <- 
  factor(predict_GER_table_sensAnalysis_phase1_permth$MC_variable, 
         levels=c('all','P_cost','Es','E_FnU'))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables", "P", "Es", "E F+U")
names(MC_var_factor.labs) <- c("all", "P_cost", "Es","E_FnU")

plot_predict_GER_sensAnalysis_phase1_permth <- 
  predict_GER_table_sensAnalysis_phase1_permth %>%
  dplyr::filter(sex == "N/A") %>% 
  ggplot(aes(x = age_mth, y = mean_GER)) +
  geom_errorbar(aes(x = age_mth-0.5, 
                    ymin = mean_GER-GER_sd, 
                    ymax = mean_GER+GER_sd), 
                width=0, size = 0.5,
                linetype = 1,
                color="gray80") + 
  geom_point(aes(x = age_mth-0.5, y= mean_GER), 
             shape = 21, fill = "black")+
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  xlab("Age (months)") +
  ylab(bquote('Gross Energy Reqs (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5),
                     limits = c(0, 12.5)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 7),
                     limits = c(0,1100)) + 
  theme_bw()+
  theme(panel.spacing = unit(0.6, "lines"))+
  theme(panel.grid = element_blank())+
  theme(panel.spacing = unit(0.6, "lines"))+
  theme(strip.background = element_rect(
    color="transparent", fill="transparent"))+
  theme(axis.text = element_text(size = rel(1),
                                 colour = "black")) +
  theme(axis.title.x = element_text(size = rel(1),
                                    colour = "black")) +
  theme(axis.title.y = element_text(size = rel(1),
                                    colour = "black"))


plot_predict_GER_sensAnalysis_phase1_permth

```

Figure 14. Results for the sensitivity analysis on gross energy requirements (*GER*) for grey whale calves.

#### Grey whale juveniles / adults

```{r predict_GER_table_sensAnalysis_phase2,  echo= FALSE, results = "hide", message = FALSE, warning=FALSE}
predict_GER_table_sensAnalysis_phase2 <- read_csv("data/predict_GER_table_sensAnalysis_phase2.csv")

predict_GER_table_sensAnalysis_phase2 %>% 
  filter(sex=="N/A" & phase == "2" & age_yrs >= 1 )%>% 
  head() %>% kable()
kable(predict_GER_table_sensAnalysis_phase2)

predict_GER_table_sensAnalysis_phase2$MC_var_factor <- 
  factor(predict_GER_table_sensAnalysis_phase2$MC_variable, 
         levels=c('all','P_cost','Es','E_FnU'))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables", "P", "Es", "E F+U")
names(MC_var_factor.labs) <- c("all", "P_cost", "Es","E_FnU")


plot_predict_GER_table_sensAnalysis_phase2 <- predict_GER_table_sensAnalysis_phase2 %>%
  dplyr::filter(sex=="N/A" & phase == "2" & 
                  age_yrs >= 1 & age_yrs <= 30) %>% 
  ggplot() +
  # geom_line(aes(x=age_yrs, y=mean_GER),
  #           color='black',
  #           size = 1.2,
  #           linetype = "longdash") +
  
  geom_errorbar(aes(x = age_yrs-0.5, 
                    ymin = mean_GER-GER_sd, 
                    ymax = mean_GER+GER_sd), 
                width=0, size = 0.5, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_yrs-0.5, y= mean_GER), shape = 21, fill = "black")+
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  xlab("Age (years)") +
  ylab(bquote('Gross Energy Reqs (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), 
                     limits = c(0, 30.5), expand = c(0,0)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),
                     #limits = c(-50, 4000), expand = c(0,0),
                     labels = comma)+
  #ggtitle("Juveniles / non-reproductive adults")+
  
  theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(panel.spacing = unit(0.6, "lines"))+
   theme(strip.background = element_rect(color="transparent", fill="transparent"))+
  theme(axis.title = element_text(size = rel(1), colour = "black"))+
  theme(axis.text = element_text(size = rel(1), colour = "black"))
  # annotate("text", x = 28, y = 3500,
  #          label = paste0("Annual"),
  #          size = 6,
  #          hjust = 1) 

plot_predict_GER_table_sensAnalysis_phase2

```

Figure 15. Results for the sensitivity analysis on gross energy requirements (*GER*) for juveniles and non-reproductive adults.

#### Pregnant grey whales

```{r predict_GER_table_sensAnalysis_preg, echo= FALSE, results = "hide", message = FALSE, warning=FALSE, fig.width=10,fig.height=5}
#Read in data
predict_GER_table_sensAnalysis_preg <-
  as_tibble(read_csv("data/predict_GER_table_sensAnalysis_preg.csv"))


predict_GER_table_sensAnalysis_preg$MC_var_factor <- 
  factor(predict_GER_table_sensAnalysis_preg$MC_variable, 
         levels=c("all","P_cost","Es", "E_FnU", "Hg", "P_nb", "GER_calf"))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables","P", "Es", "E F+U", "Hg", "P newborn", "GER calf 6mo.")
names(MC_var_factor.labs) <- c("all","P_cost", "Es", "E_FnU", "Hg", "P_nb", "GER_calf")

plot_predict_GER_sensAnalysis_preg <- predict_GER_table_sensAnalysis_preg %>%
  #dplyr::filter(sex=="Female") %>% 
  ggplot() +
  geom_errorbar(aes(x = age_yrs, 
                    ymin = mean_GER - GER_sd, 
                    ymax = mean_GER + GER_sd), 
                width=0, size = 0.5, linetype = 1,color="gray80") + 
  geom_point(aes(x = age_yrs, y= mean_GER), 
             shape = 21, fill = "white")+
  geom_errorbar(aes(x = age_yrs, 
                    ymin = GER_foraging - sd_foraging, 
                    ymax = GER_foraging + sd_foraging), 
                width=0, size =0.5, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_yrs, y= GER_foraging), 
             shape = 21, fill = "black")+
  #facet_grid(~MC_variable)+
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  #facet_wrap(~MC_variable)+
  xlab("Age (years)") +
   ylab(bquote('Gross Energy Reqs (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), 
                     limits = c(8, 40)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4)
                     #limits = c(0,max(plot_predict_GER_sensAnalysis_preg$quant975_foraging))
  ) +
  #ggtitle("Pregnant females")+
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text = element_text(size = rel(1),
                                 colour = "black"))+
  theme(axis.title.x = element_text(size = rel(1)),
        axis.title.y = element_text(size = rel(1)))+
  theme(strip.background = element_rect(
    color="transparent", fill="transparent"))+
  theme(strip.text = element_text(size = rel(1)))

plot_predict_GER_sensAnalysis_preg              


```

Figure 16. Results for the sensitivity analysis on gross energy requirements (*GER*) for pregnant females. GER calf represents the energy required for the first 6 months.

#### Lactating grey whales

```{r predict_GER_table_sensAnalysis_lact, echo= FALSE, results = "hide", message = FALSE, warning=FALSE, fig.width=10,fig.height=5}

#Read in data
predict_GER_table_sensAnalysis_lact <-
  as_tibble(read_csv("data/predict_GER_table_sensAnalysis_lact.csv"))


predict_GER_table_sensAnalysis_lact$MC_var_factor <- 
  factor(predict_GER_table_sensAnalysis_lact$MC_variable, 
         levels=c("all","P_cost","Es", "E_FnU","GERcalf"))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables","P", "Es", "E F+U", "GER calf 3.6mo.")
names(MC_var_factor.labs) <- c("all","P_cost", "Es", "E_FnU", "GERcalf")

plot_predict_GER_sensAnalysis_lact <- predict_GER_table_sensAnalysis_lact %>%
  #dplyr::filter(sex=="Female") %>% 
  ggplot() +
  geom_errorbar(aes(x = age_yrs, 
                    ymin = mean_GER - GER_sd, 
                    ymax = mean_GER + GER_sd), 
                width=0, size = 0.5, linetype = 1,color="gray80") + 
  geom_point(aes(x = age_yrs, y= mean_GER), 
             shape = 21, fill = "white")+
  geom_errorbar(aes(x = age_yrs, 
                    ymin = GER_foraging - sd_foraging, 
                    ymax = GER_foraging + sd_foraging), 
                width=0, size = 0.5, linetype = 1, color="gray80") + 
  geom_point(aes(x = age_yrs, y= GER_foraging), 
             shape = 21, fill = "black")+
  #facet_grid(~MC_variable)+
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  #facet_wrap(~MC_variable)+
  xlab("Age (years)") +
  ylab(bquote('GER (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), 
                     limits = c(8, 40)) +  # max x-axis 30 yrs. 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4)
                     #limits = c(0,max(plot_predict_GER_sensAnalysis_lact$quant975_foraging))
  ) +
#    ggtitle("Lactating females")+
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text = element_text(size = rel(1),
                                 colour = "black"))+
  theme(axis.title.x = element_text(size = rel(1)),
        axis.title.y = element_text(size = rel(1)))+
  theme(strip.background = element_rect(
    color="transparent", fill="transparent"))+
  theme(strip.text = element_text(size = rel(1)))

plot_predict_GER_sensAnalysis_lact          

```

Figure 17. Results for the sensitivity analysis on gross energy requirements (*GER*) for lactating females. GER calf represents the energy required for the last 3.6 months of nursing while mother and calf are in the foraging grounds.


### Total energetic expenditure (*E~s~*)

#### Grey whale calves

```{r Es_sensAnalysis_phase1_permth, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}


Es_sensAnalysis_phase1_permth <-  read_csv("data/Es_sensAnalysis_phase1_permth_source_bpm.csv")

Es_sensAnalysis_phase1_permth$MC_var_factor <- 
  factor(Es_sensAnalysis_phase1_permth$MC_variable, 
         levels=c("all","pctO2", "Rs", "Vt" ))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables","%O2 absorbed", "Resp. rates", "Tidal Vol.")
names(MC_var_factor.labs) <- c("all","pctO2", "Rs", "Vt")


plot_Es_sensAnalysis_phase1_permth <- Es_sensAnalysis_phase1_permth %>% 
  #dplyr::filter(age_yrs == i) %>% 
  ggplot(aes(x = age_mth-0.5, y =Es)) +
  geom_errorbar(aes(ymin = Es - Es_sd, ymax = Es + Es_sd),
                linetype = 1, size = 0.5, colour = 'gray80', width = 0) +
  geom_point() +
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  xlab("Age (months)") +
  ylab(bquote('Energetic expenditure (MJ day'^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), 
                      limits = c(0, 12.5)) +  # max x-axis 30 yrs. 
  scale_y_continuous(label = comma, 
                     breaks = scales::pretty_breaks(n = 10)
                     #limits = c(0,1500)
                     ) +
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(panel.spacing = unit(0.6, "lines"))+
  theme(strip.background = element_rect(
    color="transparent", fill="transparent"))+
  theme(axis.title.x = element_text(size = rel(1),
                                    colour = "black"))+
  theme(axis.title.y = element_text(size = rel(1),
                                    colour = "black"))+
  theme(axis.text = element_text(size = rel(1),
                                 colour = "black"))  
#ggtitle("Es table Phase 1 - per mth")

plot_Es_sensAnalysis_phase1_permth

```

Figure 18. Results for the sensitivity analysis on total energetic expenditure (*E~s~*) for grey whale calves.

#### Grey whale juveniles and adults, including pregnant and lactating females

```{r Es_sensAnalysis_alladults_preg_lact, echo= FALSE, results = "hide", message = FALSE, warning=FALSE}

#Read in data
Es_sensAnalysis_alladults_preg_lact_peryear <-
  as_tibble(read_csv("data/Es_sensAnalysis_alladults_preg_lact_peryear_source_bpm.csv"))


# apply factors to enable facet wrap 
Es_sensAnalysis_alladults_preg_lact_peryear$Lifestage_f =
  factor(Es_sensAnalysis_alladults_preg_lact_peryear$Lifestage,  
         levels=c('Juvenile/Adult','Pregnant','Lactating'))


Es_sensAnalysis_alladults_preg_lact_peryear$MC_var_factor <- 
  factor(Es_sensAnalysis_alladults_preg_lact_peryear$MC_variable, 
         levels=c("all","pctO2","Rs", "Vt"))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables","%O2 absorbed", "Resp. rates", "Tidal Vol.")
names(MC_var_factor.labs) <- c("all","pctO2","Rs", "Vt")

# Designing the plot 
plot_Es_sensAnalysis_alladults_preg_lact_yearly_perday <-
  Es_sensAnalysis_alladults_preg_lact_peryear %>% 
  ggplot(aes(x = age_yrs, y =Es_perday)) + #shape = Lifestage
  #facet_grid(Lifestage_f ~ MC_variable)+
    facet_grid(Lifestage_f~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  geom_errorbar(aes(x = age_yrs, ymin = Es_perday - Es_sd_perday, ymax = Es_perday + Es_sd_perday),
                size = 0.5, colour = 'gray80', width = 0, linetype = 1) +
  # geom_ribbon(aes(x = age_yrs, ymin = Es - Es_sd, ymax = Es + Es_sd),
  #                 fill = 'gray40', alpha = 0.3) +
  geom_point(size = 1)+
  #geom_line()   +
  xlab("Age (years)") +
  
  ylab(bquote('Energetic expenditure (MJ day'^'-1'*')')) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5), 
                     limits = c(0, 32)) +  # max x-axis 30 yrs. 
  scale_y_continuous(label = comma, 
                     breaks = scales::pretty_breaks(n = 6),
                     limits = c(0,1500))+
  # scale_y_continuous(label = function(x){(x/1000)}, breaks = scales::pretty_breaks(n = 8),
  #                    limits = c(0, 1000000))+
  
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text = element_text(size = rel(1),
                                 colour = "black"))+
  theme(axis.title.x = element_text(size = rel(1)),
        axis.title.y = element_text(size = rel(1)))+
  theme(strip.background = element_rect(
    color="transparent", fill="transparent"))+
  theme(panel.spacing = unit(0.6, "lines"))+
  theme(strip.text = element_text(size = rel(1)))
#theme(legend.position = "none")

plot_Es_sensAnalysis_alladults_preg_lact_yearly_perday


```

Figure 19. Results for the sensitivity analysis on *E~s~* (total energetic expenditure) for Juvenile and adult grey whales, including pregnant and lactating females.







```{r plots_Es_phase2_peryear, echo=FALSE}



plot_Es_sensAnalysis_phase2_peryear <- Es_sensAnalysis_phase2_peryear %>% 
  dplyr::filter(Lifestage == "Juvenile/Adult" & age_yrs >= 1 & age_yrs < 30) %>% 
  ggplot() +
  geom_errorbar(aes(x = age_yrs, ymin = Es_perday - Es_sd_perday, 
                    ymax = Es_perday + Es_sd_perday),
                linetype = 3, colour = 'gray40', width = 0) +
  geom_point(aes(x = age_yrs, y =Es_perday)) +
  facet_grid(~MC_variable)+
  xlab("Age (years)") +
  ylab(bquote('Energetic expenditure (MJ day '^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
                     #limits = c(0, 30.5), expand=c(0,0)) +  # max x-axis 30 yrs.
  scale_y_continuous(label=comma,
                     breaks = scales::pretty_breaks(n = 10))+
                     #limits = c(0,1400), expand=c(0,0))+
  theme_bw()+
theme_bw() +
  theme(panel.grid = element_blank())+
  theme(legend.position = 0)+
  # theme(plot.background = element_rect(fill = "black"))+
  # theme(panel.background = element_rect(fill = "black"))+
  theme(panel.border = element_blank())+
  theme(axis.line = element_line(size = 1, colour = "gray75"))+
  theme(axis.text = element_text(colour = "black", size = rel(1.2)))+
  theme(axis.title.y = element_text(colour = "black", 
                                    size = rel(1.4), angle = 90))+
  theme(axis.title.x = element_text(colour = "black", 
                                    size = rel(1.4)))

plot_Es_sensAnalysis_phase2_peryear

```

Es for all life stages including pregnant and lactating

```{r Es preg lact, echo=FALSE}

#read in preg and lact table. 
Es_preg_table  <-  read_csv("data/Es_sensAnalysis_preg_peryear_source_bpm.csv")
Es_lact_table <- read_csv("data/Es_sensAnalysis_lact_peryear_source_bpm.csv")


#Limit tibbles to 30 yrs
Es_lact_table_max30 <-   Es_lact_table %>% dplyr::filter(age_yrs <= 30)
Es_preg_table_max30 <-   Es_preg_table %>% dplyr::filter(age_yrs <= 30)

Es_sensAnalysis_alladults_preg_lact_peryear <-
  Es_sensAnalysis_phase2_peryear %>%
  dplyr::filter(Lifestage == "Juvenile/Adult" & age_yrs >= 1 & age_yrs <= 30)

Es_sensAnalysis_alladults_preg_lact_peryear <-
  rbind(Es_sensAnalysis_alladults_preg_lact_peryear,
        Es_preg_table_max30,
        Es_lact_table_max30)

Es_sensAnalysis_alladults_preg_lact_peryear %>% write_csv("data/Es_sensAnalysis_alladults_preg_lact_peryear_source_bpm.csv",
                                                          na = "",
                                                          append = FALSE)


# apply factors to enable facet wrap
Es_sensAnalysis_alladults_preg_lact_peryear$Lifestage_f =
  factor(Es_sensAnalysis_alladults_preg_lact_peryear$Lifestage,
         levels=c('Juvenile/Adult','Pregnant','Lactating'))



# plots

plot_Es_sensAnalysis_alladults_preg_lact_yearly_perday <- Es_sensAnalysis_alladults_preg_lact_peryear %>%
  ggplot(aes(x = age_yrs, y =Es_perday)) + #shape = Lifestage
  facet_grid(Lifestage_f ~ MC_variable)+
  geom_errorbar(aes(x = age_yrs, ymin = Es_perday - Es_sd_perday, ymax = Es_perday + Es_sd_perday),
                colour = 'gray40', width = 0, linetype = 3) +
  # geom_ribbon(aes(x = age_yrs, ymin = Es - Es_sd, ymax = Es + Es_sd),
  #                 fill = 'gray40', alpha = 0.3) +
  geom_point()+
  #geom_line()   +
  xlab("Age (years)") +

  ylab(bquote('Energetic expenditure (MJ day'^'-1'*')')) +

  scale_x_continuous(breaks = scales::pretty_breaks(n = 5),
                     limits = c(0, 32)) +  # max x-axis 30 yrs.
  scale_y_continuous(label = comma,
                     breaks = scales::pretty_breaks(n = 6),
                     limits = c(0,1500))+
  # scale_y_continuous(label = function(x){(x/1000)}, breaks = scales::pretty_breaks(n = 8),
  #                    limits = c(0, 1000000))+

  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text = element_text(size = rel(1),
                                 colour = "black"))+
  theme(axis.title.x = element_text(size = rel(1)),
        axis.title.y = element_text(size = rel(1)))+
  theme(strip.background =element_rect(fill="transparent",
                                       colour = "transparent"))+
  theme(strip.text = element_text(size = rel(1)))
#theme(legend.position = "none")

plot_Es_sensAnalysis_alladults_preg_lact_yearly_perday




```

