---
title: "Energy Expenditure (Es) Sensitivity Analysis - Phase 1"
author: "Selina Agbayani"
date: "02 March 2022 - code updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  html_document:
  
editor_options: 
  chunk_output_type: console
---

NOTE - this code will run for a long time.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)

## Setup

# Initialize all libraries, set paths for output figures, then import the gw_observations.csv
# install.packages("tidyr")
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("ggplot2")
# install.packages("knitr")
# install.packages("nls2")
# install.packages("nlstools")
# install.packages("nlme")
# install.packages("minpack.lm")
# install.packages("scales")
# install.packages("ggpmisc")
# install.packages("extrafont")
#install.packages("truncnorm")
#install.packages("grid")


library(tidyverse)
# library(tidyr)
# library(dplyr)
# library(ggplot2)
library(knitr)
library(nls2)
library(nlstools)
library(nlme)
library(minpack.lm)
library(scales)
library(ggpmisc)
library(reshape2)
library(extrafont)
library(truncnorm)
library(grid)


# #quiet col types messages
# options(readr.show_col_types = FALSE)



```


```{r functions, include=FALSE}

#Call R script with custom functions
source(paste0(getwd(),"/functions.R"), local = knitr::knit_global())
```   


```{r setpaths}
############ Set path for output figures: ###############
Figurespath <- paste0(getwd(), "/Es/figures", collapse = NULL)
Figurespath
############ Set path for input & output data  ###########
datapath <- paste0(getwd(), "/data", collapse = NULL) 
datapath




```

```{r readindata}
# Read in Tidal Volume table - Vt
Vt_table_phase1 <- as_tibble(read_csv("data/Vt_table_phase1.csv"),                                
                          col_types = (list(cols(age_yrs = col_double(),
                                                 Vt_mean = col_double(),
                                                 Vt_sd = col_double(),
                                                 quant025 = col_double(),
                                                 quant975 = col_double()
                                                 )
                                            )
                                       )
                          )


# Read in Activity Cost Reference data from csv  ORIGINAL SOURCE VALUES
A_cost_reference <- as_tibble(
  read_csv("data/ActivityCost_ReferenceData_BreathsPerDay_Table_VA_2017_original_sources.csv"),
  col_types = (list(cols(  ID = col_double(),
                           Lifestage = col_character(),
                           Description = col_character(),
                           Activity_stages = col_character(),
                           no_days = col_double(),
                           source_no_days = col_character(),
                           bpm = col_double(),
                           se_bpm = col_double(),
                           source_bpm = col_character(),
                           age_yrs = col_double(),
                           age_yrs_min = col_double(),
                           age_yrs_max = col_double(),
                           pct_O2 = col_double(),
                           pct_O2_sd = col_double()
  )
  )
  )
)

kable(head(A_cost_reference))

age_yr_tibble <- as_tibble(
  read_csv("data/age_yr_tibble.csv"), 
  col_types = (list(ID = col_integer(),
                    month = col_character(),
                    no_days_in_mth = col_double(),
                    age_mth = col_double(),
                    no_days_cumul = col_double(),
                    age_yrs = col_double()
  )
  )
)



```

Total metabolic energy expenditure at a given stage (E~s~) 
------------------------------------------------------------------------

E~s~ = 0.02 x %O~2~ x T~s~ x R~s~ x V~t~ - Sumich (1986)

where:  

0.02 - Amount of heat produced in MJ/L O2 consumed (Kleiber 1961)  
%O~2~ - Extraction efficiency per breath  
T~s~ - The no. of days in that stage  
R~s~ - Respiration rate (breaths/day)  
V~t~ - Tidal lung volume (L)  



```{r Es_phase1_sensAnalysis}

Lifestage <-  "Calf"

#Pull respiration rate data for calves from Activity Cost Reference (A_cost_Reference) table 
A_cost_calf <-  A_cost_reference %>%
  select(Lifestage, Activity_stages, age_yrs, no_days, bpm, se_bpm) %>%
  filter(Lifestage == "Calf")  


kable(A_cost_calf)

# Calves 
Es_sensAnalysis_phase1 <- as.data.frame(matrix(ncol = 18, nrow = 0))

colnames(Es_sensAnalysis_phase1) <- c("age_yrs","Lifestage", 
                               "Activity_stages", "no_days", 
                               "MC_variable", 
                               "mean_bpm", "se_bpm", "mean_bpd",
                               "Vt_mean", "Vt_sd",
                               "Es_perday", "Es_perday_sd",
                               "Es_perday_quant025", "Es_perday_quant975",
                               #2.5% and 97.5% quantile 
                               #from bootstrap estimates 
                               "Es","Es_sd","Es_quant025","Es_quant975"
)            




Es_sensAnalysis_phase1 <- 
  as_tibble(Es_sensAnalysis_phase1, 
            col_types = (list(ID = col_integer(), 
                              age_yrs = col_double(),
                              Lifestage = col_double(),
                              Activity_stages = col_character(),
                              no_days = col_double(),
                              MC_variable = col_character(),
                              mean_bpm = col_double(),
                              mean_bpd = col_double(),
                              se_bpm = col_double(),
                              Vt_mean = col_double(),
                              Vt_sd = col_double(), 
                              Es_perday = col_double(),
                              Es_perday_sd  = col_double(),
                              Es_perday_quant025 = col_double(),
                              Es_perday_quant975 = col_double(),
                              Es = col_double(),
                              Es_sd = col_double(),
                              Es_quant025 = col_double(),
                              Es_quant975 = col_double()
            )
            )
  )



#Original code was run with MC_reps <- 10000  and took a very long time
#To test and explore the code, use less reps 

MC_reps = 10000

for (a in seq(from = 1, to = 12, by = 1)){
  for (MC_var in c("all","Rs", "Vt", "pctO2")){
    
    age_yrs_i <- age_yr_tibble %>% 
     filter(age_mth == a)  %>% 
      pull(age_yrs)
    
    age_mid_i <- age_yr_tibble %>% 
      filter(age_mth == a-0.5) %>% 
      pull(age_yrs)
    
    #Lifestage <-  "Calf"
    
    strcolname <- as.character(age_yrs_i)
    A_cost_calf_i <- filter(A_cost_calf, 
                                   round(A_cost_calf$age_yrs*12) == a)
    
    if (nrow(A_cost_calf_i) == 0) {
      row <- NA
    }else{
      
      Activity_stages <- A_cost_calf_i$Activity_stages
      
      for (act in Activity_stages){
        activity <- act
        Lifestage
        activity
        
        strcolname <- as.character(age_yrs_i)
        
        A_cost_i <- A_cost_reference %>% dplyr::filter(
          Activity_stages == activity & 
            Lifestage == "Calf" 
          & round(age_yrs, 3) == round(age_yrs_i,3))
        
        A_cost_i
        
        # Lifestage <- A_cost_i$Lifestage
        # A_stage <- a
        
        set.seed(12345)
        
        
        #plot(pct_O2)
        #No. of days at each age (in months)
        no_days_i <-  A_cost_i$no_days   
        
        #Respiration rates at each activity stage 
        bpm_i <- A_cost_i$bpm 
        bpm_sd_i <- A_cost_i$se_bpm 
        
        #log() computes the natural logarithms (Ln)
        meanlog_bpm_i <- log(bpm_i^2 / sqrt(bpm_sd_i^2 + bpm_i^2))  #location
        meanlog_bpm_sd_i <- sqrt(log(1 + (bpm_sd_i^2 / bpm_i^2)))  #shape
        
        bpm_tibble <- as.data.frame(matrix(ncol = 2, nrow=0))
        colnames(bpm_tibble) <- c("mean_bpm", "sd_bpm")
        bpm_tibble <- as_tibble(bpm_tibble,
                                col_types = (list(ID = col_integer(), 
                                                  mean_bpm = col_double(),
                                                  sd_bpm = col_double()
                                                  
                                )))
        
        
        for (i in seq(from = 1, to = MC_reps, by = 1)){
          
          #generate lognormal distribution of breaths per minute (bpm)
          bpm <- rlnorm(MC_reps, meanlog_bpm_i, meanlog_bpm_sd_i)  
          # draws <- rlnorm(n=1000000, location, shape)
          # https://msalganik.wordpress.com/2017/01/21/making-sense-of-the-rlnorm-function-in-r/
          # plot(bpm)
          mean_bpm <- mean(bpm)
          sd_bpm <- sd(bpm)
          bpm_row <- tibble(mean_bpm = mean_bpm,
                            sd_bpm = sd_bpm)
          bpm_tibble <- rbind(bpm_tibble, bpm_row) 
          bpm_row <- NA
        }
        
        
        mean_bpm_i <- mean(bpm_tibble$mean_bpm)
        se_bpm_i <- sd(bpm_tibble$mean_bpm)     #sd of mean of means
        
        mean_bpm_reps <- bpm_tibble$mean_bpm
        

        # breaths/day
        bpd <- mean_bpm_reps * 60 * 24
        # plot(bpd)
        bpd_mean_i <-  mean(bpd)
        
        if (MC_var == "all" || MC_var == "Rs"){
          bpd_sd_i <- sd(bpd)  
        } else {
          bpd_sd_i = 0
        }
        
        
        #pull in tidal volume estimates (mass dependent)
        Vt_table_i <- filter(
          Vt_table_phase1, 
          round(Vt_table_phase1$age_yrs,digits = 3) == round(age_mid_i,digits =3)
        )
        
        Vt_i <- Vt_table_i$Vt_mean
        
        if (MC_var == "all" || MC_var == "Vt"){
          Vt_sd_i <- Vt_table_i$Vt_sd  
        } else {
          Vt_sd_i = 0
        }
        
        set.seed(12345)
        #generate normal distribution of Vt estimates
        Vt <- rnorm(MC_reps, Vt_i, Vt_sd_i)
        Vt_mean_i <- mean(Vt)
        Vt_sd_i <- sd(Vt)
        #plot(Vt)
        
        # O2 extraction efficiency
        if (a <= 5) {
          pct_O2_i <- 10.5/100
          
          if (MC_var == "all" || MC_var == "pctO2"){
            pct_O2_sd_i <- 3/100  
          } else {
            pct_O2_sd_i = 0
          }
          
        } else if (a > 5){
          pct_O2_i <- 11/100
          
          if (MC_var == "all" || MC_var == "pctO2"){
            pct_O2_sd_i <- 2.7/100  
          } else {
            pct_O2_sd_i = 0
          }
          
        }
        
        
        #generate normal distribution of pct_O2
        pct_O2 <- rnorm(MC_reps, pct_O2_i, pct_O2_sd_i)  
        
        
        #Energetic expenditure per day
        Es_perday <- 0.02 * pct_O2* bpd * Vt   
        #plot(Es_perday)
        
        Es_perday_mean <- mean(Es_perday)
        Es_perday_sd <- sd(Es_perday)
        Es_perday_quant025 <- quantile(Es_perday, 0.025, na.rm = TRUE)
        Es_perday_quant975 <- quantile(Es_perday, 0.975, na.rm = TRUE)
        
        #Energetic expenditure per month
        Es = Es_perday * no_days_i
        #plot(Es)
        Es_mean = mean(Es)
        Es_sd = sd(Es)
        Es_quant025 <- quantile(Es, 0.025, na.rm = TRUE)
        Es_quant975 <- quantile(Es, 0.975, na.rm = TRUE)
        
        
        # "age_yrs", "Es", "Es_sd", "quant025", "quant975" 
        row <- tibble(age_yrs = age_yrs_i,
                      Lifestage = Lifestage,
                      Activity_stages = activity,
                      no_days = no_days_i,
                      MC_variable = MC_var,
                      mean_bpm = mean_bpm_i,
                      se_bpm = se_bpm_i,
                      mean_bpd = bpd_mean_i,
                      Vt_mean = Vt_mean_i,
                      Vt_sd = Vt_sd_i,
                      Es_perday = Es_perday_mean, 
                      Es_perday_sd = Es_perday_sd, 
                      Es_perday_quant025 = Es_perday_quant025,
                      Es_perday_quant975  = Es_perday_quant975,
                      Es = Es_mean, 
                      Es_sd = Es_sd,
                      Es_quant025 = Es_quant025, 
                      Es_quant975 = Es_quant975)
        
        # append new row of data to Es_sensAnalysis_phase1
        Es_sensAnalysis_phase1 <- rbind(Es_sensAnalysis_phase1, row)
        row <-  NA #reset row for next iteration
        
      } 
    }
  }
}




kable(head(Es_sensAnalysis_phase1))

#save Es_sensAnalysis_phase1 table
Es_sensAnalysis_phase1 %>% write_csv("data/Es_sensAnalysis_phase1_source_bpm.csv", na = "", append = FALSE)

```


```{r Es_phase1_permth}
Es_sensAnalysis_phase1$age_mth <- 
  round(Es_sensAnalysis_phase1$age_yrs *12)

Es_sensAnalysis_phase1_permth <- Es_sensAnalysis_phase1


#pull out blank Es_subtable
Es_subtable <- Es_sensAnalysis_phase1 %>% 
  dplyr::filter(age_yrs >999, 
                Lifestage == Lifestage)
kable(Es_subtable)


Es_sensAnalysis_phase1_permth  <- Es_subtable %>%  
  dplyr::select(age_yrs, age_mth, Lifestage, no_days, 
                MC_variable, Es, Es_sd, Es_perday, Es_perday_sd)

## Calculate sum of Es for phase 1 (per mth)
for (i in seq(from = 1, to = 12, by = 1)){
  for (MC_var in c("all","Rs", "Vt", "pctO2")){
    
    age <- age_yr_tibble %>% dplyr::filter(age_mth == i) %>%  pull(age_yrs)
    age_mth <-  i
    
    Es_subtable <- Es_sensAnalysis_phase1 %>% 
      dplyr::filter(age_yrs == age & MC_variable == MC_var)
    no_days <- sum(Es_subtable$no_days)
    Es <- sum(Es_subtable$Es)
    
    sum_of_variances <- 0
    for (row in 1:nrow(Es_subtable)){
     
      Es_subtable_i <- Es_subtable[row, "Es"]
      Es_sd_i <- Es_subtable[row, "Es_sd"]
      sum_of_variances <- sum_of_variances + (Es_sd_i)^2
    }
    Es_sd  <-  sqrt(sum_of_variances$Es_sd)
    
    Es_perday <- Es / no_days
    Es_perday_sd <- Es_sd / no_days
    
    newRow <- tibble(age_yrs = age,
                     age_mth = i,
                     Lifestage = Lifestage,
                     no_days = no_days,
                     MC_variable = MC_var,
                     Es = Es, 
                     Es_sd = Es_sd,
                     Es_perday = Es_perday,
                     Es_perday_sd = Es_perday_sd
    )
    
    Es_sensAnalysis_phase1_permth <- rbind(Es_sensAnalysis_phase1_permth, newRow)
  }
}



kable(head(Es_sensAnalysis_phase1_permth))



Es_sensAnalysis_phase1_permth %>% write_csv("data/Es_sensAnalysis_phase1_permth_source_bpm.csv", na = "", append = FALSE)

```

E~s~ per month - Phase 1 - stacked plot from sensitivity analysis


```{r plots_Es_phase1_stacked}
Es_sensAnalysis_phase1 <- read_csv("data/Es_sensAnalysis_phase1_source_bpm.csv")

plot_Es_sensAnalysis_phase1 <- Es_sensAnalysis_phase1 %>% 
  ggplot(aes(x = age_yrs, y = Es))+
  geom_errorbar(aes(ymin = Es - Es_sd, ymax = Es + Es_sd),
                colour = "blue", width = 0.02)+
  geom_point()+
  facet_grid(MC_variable ~ Activity_stages,
             labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
  ggtitle("Es table - phase 1")

plot_Es_sensAnalysis_phase1


```


E~s~ per month - Phase 1 - plot from sensitivity analysis

```{r plots_Es_phase1_permth, echo=FALSE}
Es_sensAnalysis_phase1_permth$MC_var_factor <- 
  factor(Es_sensAnalysis_phase1_permth$MC_variable, 
         levels=c("all","Rs", "Vt", "pctO2"))

# New facet label names for MC_var_factor variable
MC_var_factor.labs <- c("All variables","Rs", "Vt", "% O2")
names(MC_var_factor.labs) <- c("all","Rs", "Vt", "pctO2")


plot_Es_sensAnalysis_phase1_permth <- Es_sensAnalysis_phase1_permth %>% 
  ggplot(aes(x = age_mth-0.5, y =Es)) +
  geom_errorbar(aes(ymin = Es - Es_sd, ymax = Es + Es_sd),
                linetype = 3, colour = 'gray40', width = 0) +
  geom_point() +
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  xlab("Age (months)") +
  ylab(bquote('Energetic expenditure (MJ)')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 11))+ 
  scale_y_continuous(label = comma, 
                     breaks = scales::pretty_breaks(n = 10)
                     
                     ) +
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(panel.spacing = unit(1, "lines"))+
  theme(strip.background = element_rect(
    color="black", fill="transparent"))+
  theme(axis.title.x = element_text(size = rel(1.4),
                                    colour = "black"))+
  theme(axis.title.y = element_text(size = rel(1.4),
                                    colour = "black"))+
  theme(axis.text = element_text(size = rel(1.0),
                                 colour = "black"))  

plot_Es_sensAnalysis_phase1_permth



```

E~s~ per day at each age (month) - Phase 1 - plot from sensitivity analysis

```{r plots_Es_phase1_mthly_perday, echo=FALSE}


plot_Es_sensAnalysis_phase1_mthly_perday <- Es_sensAnalysis_phase1_permth %>% 
    ggplot(aes(x = age_mth-0.5, y =Es_perday)) +
  geom_errorbar(aes(ymin = Es_perday - Es_perday_sd, ymax = Es_perday + Es_perday_sd),
                linetype = 3, colour = 'gray40', width = 0) +
  geom_point() +
  facet_grid(~MC_var_factor, 
             labeller = labeller(MC_var_factor = MC_var_factor.labs)) +
  xlab("Age (months)") +
  ylab(bquote('Energetic expenditure (MJ day'^'-1'*')')) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6))+ 
  scale_y_continuous(label = comma, 
                     breaks = scales::pretty_breaks(n = 10)
                    ) +
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(panel.spacing = unit(1, "lines"))+
  theme(strip.background = element_rect(
    color="black", fill="transparent"))+
  theme(axis.title.x = element_text(size = rel(1.4),
                                    colour = "black"))+
  theme(axis.title.y = element_text(size = rel(1.4),
                                    colour = "black"))+
  theme(axis.text = element_text(size = rel(1.0),
                                 colour = "black"))  

plot_Es_sensAnalysis_phase1_mthly_perday



```
